---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wisdotcrashdatabase

<!-- badges: start -->
<!-- badges: end -->

The goal of wisdotcrashdatabase is to make data pulls and data analysis much easier in an R environment. To set up an R environment, you must convert the entire database into CSVs. I wrote automated scripts for these. Haven package is an option, but a couple things were not working properly. This package has functions to get certain crash flags, relabel by age or injury type for easier grouping.

## Installation

This package will never be released to CRAN as it's used internally at WisDOT.

The package can be installed from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("jacciz/wisdotcrashdatabase")
```

## Setting up environment
The format data in this package must be in a .fst, while narratives are read as a .csv. In order to set the up, SAS code is run for each year and is exported to a folder of your choosing. Run the code in inst/extdata/SAS_to_FST/ to do this. Data for each year must be run. Place all exported CSV data in the same folder.

Next, run CSV_to_FST.R. This reads the CSV files and exports as a FST. This is a batch job and can be run only once. The .ps1 file automates the translation from SAS to .fst by running the command in Shell. Be sure to chamge the system path for all of this code.

## Importing data

Old and new database data can be imported into R via a single data frame. If 'years' for a new db is selected, then columns will have to be selected. Certain old db columns are renamed to match the new db. Some columns will always be selected (see Documentation). Import is done with one function:

#### Example: Number of crashes
```{r example, echo=T, message=FALSE}
library(wisdotcrashdatabase)
library(lubridate)
library(magrittr)
library(dplyr)

crash <- import_db_data(
  filepath = "C:/CSV/csv_from_sas/fst/",
  db_type = "crash",
  years_old = c("15", "16"),
  years = c("17", "18"),
  columns = c("CRSHMTH")
)
crash %>% count(year(CRSHDATE))
```

## Crash flags
There are also functions to get crash flags. When run, a new column is added with either a "Y" or "N" if that flag exists.
Flags include:

* Deer 
* Distracted driver 
* Impaired driver 
* Speed 
* Teen driver 
* Older driver_flag 
* Seatbelt_flag_by_unit 
* Suspected drug or alcohol 

#### Example: Number of speed crashes
```{r speed flag example, echo=T}

import_db_data(
  filepath = "C:/CSV/csv_from_sas/fst/",
  db_type = "person",
  years_old = c("16"),
  years = c("17", "18"),
  columns = c("STATNM", "DRVRPC")
) %>% get_driver_flags (flags = "speed") %>%
  filter(speed_flag == "Y") %>%
  distinct(CRSHNMBR, .keep_all = TRUE) %>%
  count(year = year(CRSHDATE))
```

## Relabel functions
There's also functions that may summarize data for analysis.
These include:

* Bin crash times 
* Bin age groups 
* Get county names based on cntycode 
* Bin injures into Killed, Injured, No Injury 

#### Example: People injured
```{r relabel example, echo=T}
import_db_data(
  filepath = "C:/CSV/csv_from_sas/fst/",
  db_type = "person",
  years_old = c("16"),
  years = c("17", "18")
) %>% relabel_person_variables(relabel_by = "wisinj") %>%
  count(year = year(CRSHDATE), inj)
```
## Other functions
* Get motorcyclists 
* Import crash narratives
