---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wisdotcrashdatabase

<!-- badges: start -->
<!-- badges: end -->

wisdotcrashdatabase makes data pulls and data analysis much easier inside an R environment. To set up an R environment, you must convert the entire database into FST format. I wrote automated scripts for these. Haven package is an option, but formatting of the labels was an issue. This package can gather crash flags, relabel by age or injury type for easier grouping, and can lookup county/municiaplity codes.

## Installation

This package will never be released to CRAN as it's used internally at WisDOT.

The package can be installed from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("jacciz/wisdotcrashdatabase")
```

## Setting up the environment
The format data in this package must be in a .fst, but narratives must be in a .csv. In order to set the up, SAS code is run for each year and is exported to a folder of your choosing. This is done by running the three SAS scripts in inst/extdata/SAS_to_FST/. Use the appropriate code for the appropriate crash db year. This code outputs a CSV of crash, person, vehicle, and narrative. Scripts must be run for each year. These CSVs must be in a single folder and you may have to change the folder path in the code.

Next, run the R code to translate the .csv to .fst with 'CSV_to_FST.R'. This reads all the CSV files and exports each as a FST. This is a batch job and can run only ONCE. To update current year (like 2021), the .ps1 file automates this translation from SAS to .fst by running the command in Shell. Be sure to change the file paths!

## Importing data

Old and new database data can be imported into R as a single data frame. If 'years' for a new db is selected, then columns will have to be selected. Certain old db columns are renamed to match the new db. Some columns will always be selected (see Documentation). Import is done with one function:

#### Example: Number of crashes
```{r example, echo=T, message=FALSE}
library(wisdotcrashdatabase)
library(lubridate)
library(magrittr)
library(dplyr)

crash <- import_db_data(
  filepath = "C:/CSV/csv_from_sas/fst/",
  db_type = "crash",
  years_old = c("15", "16"),
  years = c("17", "18"),
  columns = c("CRSHMTH")
)
crash %>% count(year(CRSHDATE))
```

## Crash flags functions
There are also functions to get crash flags. When run, a new column is added with either a "Y" or "N" if that flag exists.
Flags include:

* Deer 
* Distracted driver 
* Impaired driver 
* Speed 
* Teen driver 
* Older driver_flag 
* Seatbelt_flag_by_unit 
* Suspected drug or alcohol 

#### Example: Number of speed crashes
```{r speed flag example, echo=T}

import_db_data(
  filepath = "C:/CSV/csv_from_sas/fst/",
  db_type = "person",
  years_old = c("16"),
  years = c("17", "18"),
  columns = c("STATNM", "DRVRPC")
) %>% get_driver_flags (flags = "speed") %>%
  filter(speed_flag == "Y") %>%
  distinct(CRSHNMBR, .keep_all = TRUE) %>%
  count(year = year(CRSHDATE))
```

## Relabel functions
There's also functions that may summarize data for analysis.
These include:

* Bin crash times 
* Bin age groups 
* Get county names based on cntycode 
* Bin injures into Killed, Injured, No Injury 

#### Example: People injured
```{r relabel example, echo=T}
import_db_data(
  filepath = "C:/CSV/csv_from_sas/fst/",
  db_type = "person",
  years_old = c("16"),
  years = c("17", "18")
) %>% relabel_person_variables(relabel_by = "wisinj") %>%
  count(year = year(CRSHDATE), inj)
```
## Other functions
* Get motorcyclists 
* Import crash narratives 
* Lookup county code and fips 
# Lookup municode by name and ctv 

## For help
Load the wisdotcrashdatabase package in R and type in the console `?import_db_data` or any other function.
